---
title: "NY Times API Shiny App"
author: "Layla Quinones"
date: "3/23/2020"
output: rmdformats::readthedown
---

```{r, warning = FALSE, echo = FALSE, include = FALSE}
library(jsonlite)
library(tidyverse)
library(kableExtra)
#Load Packages for shiny app
library(shiny)
library(DT)
#API KEY
KEY <- "oZ6c0FgayC1zkrhn3lGBM0s07BXlL9e4"
```

>Note: jsonlite, tidyverse, kabelExtra, shiny,and DT libraries used.

# Creating a Program that accesses New York Times Articles 

When I first started this project and read the assigment I thought that I would need to create a user interface for people to use the New York Times API to search for articles. With the user in mind, I decided that my programshould return a list of articles in a dataframe with abstracts and links so that people can see a breif summary of the article and use the link to read the article if they are intersted. The steps I took in developing this program is as follows:

- Select a keyword to give the api to return results: I chose **corona virus**

- Select a start date and end date to narrow your search: I chose between **March 23, 2019** and **March 23, 2020**

- Identify your API key by registering for one at [NYTimes API Website](https://developer.nytimes.com/apis)

- Use the above variables to construct the url for the api call and `paste0()` to concatinate all variables into one string 

- Query the API using `fromJSON()` and piping the resulting object into a dataframe using `data.frame()`.

- To help tidy the data frame for specifc components that I thought were helpful, I also selected only the columns with an abstract and a url - so that people who are using this code can quickly judge if they want to read a particular article.

```{r}
#Select keyword
keyword <- "corona+virus" # Need to use + to string together separate words

#Time fram e: within the last year
beginDate <- "20190323"
endDate <- "20200323"

#API Key HIDDEN
#KEY <- "######"

#Construct url
url <- paste0("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=",keyword,"&begin_date=",beginDate,"&end_date=",endDate,"&facet_filter=true&api-key=",KEY, sep="")

#Query the API 
initialQuery <- fromJSON(url, flatten = T) %>%data.frame()

#Descriptions and links only
desLink <- initialQuery %>% 
  select(response.docs.abstract, response.docs.web_url) %>%
  rename("Abstract" = response.docs.abstract, "Link"= response.docs.web_url)

#Entire dataframe
kable(initialQuery) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

#Links and abstracts
kable(desLink) %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "200px")

```

# Shiny App for User Interface

## [Click to Launch Shiny App](https://msqcsnyc.shinyapps.io/NYTAPI/)

Now that we see how simple it is to adjust a specific query for the New York Times Article API, we are ready to start thinking about the user's experience using our tool. I decided that the user would need an interface where they can enter a keyword or a date and get results back in the form of a data table. The best tool in R for user interfaces that I know of is ShinyApps using the `shiny` library. The steps to converting the code above into a ShinyApp were as follows:

- Identify API Key and place in a variable

- Create a function that will run the program shown above (note: a keyword is **NOT** identified because that is something we would like the user to do.) A simpler url to call on the json was used ( dates omitted). This function returns a dataframe that matches the keyword input.

- Define the user interface function with a text input and a button that takes in a user's keyword on the left panel,and displays the dataframe that results in the right panel.

- In the shiny app server define a reactve function that takes in the user keyword when the button is pressed and calls the first function on the user keyword. The dataframe that results is rended in a table as an output for the shiny app

- Finally, `shinyApp` is called to build the application

```{r, warning = FALSE, results = 'hide'}
#NYT KEY HIDDEN
# KEY <- "#######"

#Creates a dataframe when keyword is given
keywordDF <- function(keyword){
  #paste together url
  url <- paste0("https://api.nytimes.com/svc/search/v2/articlesearch.json?q=", keyword,"&api-key=",KEY, sep="")
  
  fromJSON(url, flatten = T) %>% 
    data.frame() %>% 
    select(response.docs.abstract, response.docs.web_url)%>%
    rename("Abstract" = response.docs.abstract, "Link"= response.docs.web_url)
  }
  
# User Interface
ui <- fluidPage(
  titlePanel("New York Times Articles"),#Title of the Shiny APP
  sidebarLayout(
    #Left side for user input and button
        sidebarPanel(("To access NYT Articles:"), 
                     textInput("keyword", "Enter one keyword:",""),
                     actionButton("search", "Lets GO!")),
        #Main Panel for Dataframe output
        mainPanel(
          h1("NYT Article Abstracts and Links"),
          DT::dataTableOutput("dataframe")
        )
    )
)

#Server
server <- function(input, output) {
  #Define a function that takes in the user input and calls the function defined above on the keyword - When the button is pressed
  articles <- reactive({
    input$search
    
    keyword <- isolate(input$keyword)
    
    keywordDF(keyword)
    })
    #Render a datatable from the datafram output
    output$dataframe <- renderDT(articles())

}

# Run the Shiny APP (commented out for Rmd Presentation)
#shinyApp(ui = ui, server = server)
```

I chose this approach because I have made similar apps in javascript, [Click here to view my javascript version of this App](https://msqcompsci.github.io/GifsnNewsApiApp/), and this week we have been thinking about our users, user interfaces and user experience when reading and using our tools.